---
title: "The Impact of Posting Restriction on User Retention"
output: word_document
---


```{r, include=FALSE}

library(ggplot2)
library(dplyr)
library(ggridges)
retention_dataset <- read.csv("/Users/Desktop/platform_retention.csv")
```
**Introduction** \n 
 In this report, we review a comprehensive analysis of the impact of posting restriction on user engagement within a social media startup company. Our dataset includes engagement metrics from 90,189 users, each randomly assigned one of two posting models, limited or unlimited. The limited model restricts the user's daily number of posts, while the unlimited offers unrestricted posting. The primary research question of this investigation is whether posting limitation improves 7-day user retention rates in comparison to unrestricted posting. To supplement this inquiry, we are also interested in exploring site visits within the first 7 days of registration between the two models, variations in user retention after the initial day, and any additional significant relationships that could inform our next phase of testing. Through this analysis, we aim to inform site improvements that increase user engagement.
  
**Analysis** \n
                         7- Day Retention 
The following chart presents a comparative graphic of user retention rate within the first 7 days following registration, contrasting unlimited and limited posting models.
```{r Side by Side Bar Plot, echo=FALSE}
filtered_data <- retention_dataset %>%
  filter((version == 'unlimited' & retention_7 == "FALSE") |
         (version == 'limited' & retention_7 == "FALSE") |
         (version == 'unlimited' & retention_7 == "TRUE") |
         (version == 'limited' & retention_7 == "TRUE"))

# Create the grouped bar plot
ggplot(filtered_data, aes(x = version, fill = retention_7)) +
  geom_bar(position = "dodge") +
  labs(x = "Posting Version", y = "Count", fill = "7-Day Retention", title = "7-Day Retention Distribution by User Posting Version") +
  theme_minimal()
```
This visualization sheds light on the fundamental metric of user engagement, providing a clear comparison of how different posting models impact the longevity of user satisfaction. Based on the stacked bar plot, we observe very subtle differences in 7-day retention rates between the two posting versions. To further investigate these observations, formal statistical testing was conducted.

A difference of proportion test, with a significance level of 0.05 resulted in the following p-value.
```{r, echo=FALSE}
table1 <- table(retention_dataset$version, retention_dataset$retention_7)
p_val <- prop.test(x = c(8122, 7872), n = c(42505, 43168), alternative = "greater", correct = FALSE)
cat("Difference in proportion test p-value:", round(p_val$p.value, 5))
```
The result of this test suggests there is a statistically significant increase in 7-day retention among those with the limited posting model compared to the unlimited model. Additional testing yielded a confidence interval that underscored the extent of the difference in values.
```{r, echo=FALSE}
result <- prop.test(x = c(8122, 7872), n = c(42505, 43168), conf.level = .90)
confidence_interval <- result$conf.int
cat("90% Confidence Interval:", round(result$conf.int[1], 5), "to", round(result$conf.int[2], 5))
```
The 90% confidence interval revealed a potential increase in 7 day retention for limited posting, ranging from 0.43% to 1.3% when compared to unlimited posting. These findings align with the side-by-side bar plot above, providing statistical support for the subtle yet significant disparities observed between the two posting models.

Additional Contributing Variables
To explore the variables that may contribute to this difference, we examined the number of site visits by day 7 as well as the user retention 1 day post registration. 

```{r Overlapping Histograms, echo=FALSE}
new_data <- retention_dataset %>%
  filter(row_number() != 37650)
new_data$site_visits_7 <- log(new_data$site_visits_7 + 1)

ggplot(new_data, aes(x = site_visits_7, fill = version)) +
  geom_histogram(alpha = 0.7, position = "identity", bins = 15) +
  labs(title = "Log-Transformed 7-Day Site Visits by\n         User Posting Version",
       x = "Site Visits Within First 7 Days",
       y = "Frequency")
```
In the graph above, we can see that the number of site visits within the first 7 days post registration are very similar between posting versions, following the same trends. To confirm this visual observation, a two sample t-test was conducted.
```{r, echo=FALSE}
limited_visits <- new_data$site_visits_7[new_data$version == "limited"]
unlimited_visits <- new_data$site_visits_7[new_data$version == "unlimited"]
t_test_result <- t.test(limited_visits, unlimited_visits)
cat("Two sample t-test p-value:", round(t_test_result$p.value, 3))
```
With a p-value of 0.23, we can confirm that there is no statistical difference between 7-day site visits and user posting model.
Our final variable relationship to explore is 1 day retention rate. Below is a stacked bar chart representing the differences by user posting model.
```{r retention 1, echo=FALSE}
filtered_data1 <- retention_dataset %>%
  filter((version == 'unlimited' & retention_1 == "FALSE") |
         (version == 'limited' & retention_1 == "FALSE") |
         (version == 'unlimited' & retention_1 == "TRUE") |
         (version == 'limited' & retention_1 == "TRUE"))

ggplot(filtered_data1, aes(x = version, fill = retention_7)) +
  geom_bar(position = "dodge") +
  labs(x = "Posting Version", y = "Count", fill = "1-Day Retention", title = "1-Day Retention Distribution by User Posting Version") +
  theme_minimal()
```
Again, we see very subtle differences in 1-day retention rate compared to posting version. This observation was confirmed with further statistical testing resulting in the p-value below.
```{r, echo=FALSE}
contingency_table <- table(retention_dataset$retention_1, retention_dataset$version)
chi_square_test <- chisq.test(contingency_table)
cat("Association test p-value:", round(chi_square_test$p.value, 3))
```
**Conclusion**
  In addressing the primary research question, we have confirmed that limiting posting significantly increases 7 day retention among users from 0.43% to 1.3%. Additional investigations into 7-day site visits and 1-day retention showed no significant variation based on posting version. Hence, within the scope of this analysis, no other relationships among the recorded variables were found. Moving forward, we recommend extending version testing duration and considering new variables like session duration and demographics for a deeper understanding of user behavior.










